Online Balanced Graph Partitioning: Optimizations for Large Clusters

### 2020-05-04 paper/disc20clustering.tex
Lower bound todos

\maciek{Better name than $S_1$? Here and in the proof}
First, we construct a ground set of size $k-1$ on one of the clusters $S_1$.
In any configuration, there exists a node that accompanies it.
We only issue requests between such accompanying (time-varying) node, and any alone node that was initially collocated with it.
Repeating such requests, almost half of all node must visit the cluster $S_1$.
In comparison, \OPT performs one migration.

@@

\begin{proof}
	\maciek{Better name than $S_1$?}
	Fix any online algorithm \ALG{}.
	Let $I(C)$ denote the cluster where nodes of ground set $C$ are located initially.
	If $C$ contains nodes that originate from different clusters, $I(C)$ is undefined.
	Throughout the construction, we maintain ground sets for each node.
	Initially all nodes are \emph{alone} (they belong to their own trivial ground sets).

	We distinguish a cluster $S_1$, where we form a ground set of size $k-1$.
	We never dismiss any constructed ground set.
	Each cluster host exactly $k$ nodes, hence in any configuration of \ALG, there exists a single alone node in $S_1$.

	We construct a ground set containing this alone node (and some other node).
	Precisely, let $x_0$  be the alone node from $S_1$ and let $y_0$ be an arbitrary node that does not reside in $S_1$.
	We construct a ground set $\{x_0, y_0\}$.
	\ALG has them split, and the adversary produces external requests until these are collocated.
	$S_1$ already contains a ground set of size $k-1$ (that it cannot split), hence \ALG must collocate $\{x_0, y_0\}$ in another cluster.

	We continue to construct ground sets containing the current alone node from $S_1$.
	To preserve a feasible partition of nodes into clusters after collocating $\{x_0, y_0\}$,
	\ALG must replace $x_0$ with another single node (that we name $x_1$).
	Then, we merge the ground sets of $x_1$ and the largest ground set $C$ s.t. and $C \neq \{x_0,y_0\}$~$I(C) = I(x_1)$.
	Continuing this, we observe a sequence $X := x_0, x_1, x_2\dots$ of alone nodes that visit $S_1$.
	(Note that subsequent ground sets are different than the first ground set $\{x_0, y_0\}$, whose nodes were initially in different clusters.)

	We claim that this procedure (of constructing ground sets) can continue as long as a perfect partition of ground sets still exists.
	%and we construct a sequence of ground sets $X := \langle \{x_0, y_0\}, \{x_1, y_2\}, \{x_2, y_2\}, \ldots \rangle$ of length $\Omega(k \cdot \ell)$.
	Consider the moment of creating a ground set $\{x_i, y_i\}$.
	If a perfect partition of ground sets still exists, then
	\ALG{} must collocate $\{x_i,y_i\}$ in a~cluster other than $S_1$.
	To this end, it replaces $x_i$ with another alone node $x_{i+1}$.
	To continue, we must guarantee the existence of an alone node that was initially collocated with $x_{i+1}$.
	\maciek{elaborate more}

	We end this procedure prematurely (before all nodes visit $S_1$) to keep \OPT's cost low.
	The sequence must end when no perfect partition of ground sets exists ---
	however, we end it when \emph{two more} ground sets are still possible to give.
	Precisely, we repeat constructing ground sets,
	until there are only two alone nodes left that originate from the same cluster.
	We denote these two nodes by $x^*, y^*$.
	Observe that at this point there are at most $\ell+1$ alone nodes left:
	otherwise, by pigeon hole principle, there would be more pairs of alone nodes that were initially in the same cluster.

>>>>

First, we construct a ground set of size $k-1$ on one of the clusters $S$.
In any configuration, there exists a node that accompanies it.
We only issue requests between such accompanying (time-varying) node, and any alone node that was initially collocated with it.
Repeating such requests, almost half of all node must visit the cluster $S_1$.
In comparison, \OPT performs one migration.

@@

\begin{proof}
	Fix any online algorithm \ALG{}.
	Let $I(C)$ denote the cluster where nodes of ground set $C$ are located initially.
	If $C$ contains nodes that originate from different clusters, $I(C)$ is undefined.
	Throughout the construction, we maintain ground sets for each node.
	Initially all nodes are \emph{alone} (they belong to their own trivial ground sets).

	We distinguish a cluster $S$, where we form a ground set of size $k-1$.
	We never dismiss any constructed ground set.
	Each cluster host exactly $k$ nodes, hence in any configuration of \ALG, there exists a single alone node in $S$.

	We construct a ground set containing this alone node (and some other node).
	Precisely, let $x_0$  be the alone node from $S_1$ and let $y_0$ be an arbitrary node that does not reside in $S$.
	We construct a ground set $\{x_0, y_0\}$.
	\ALG has them split, and the adversary produces external requests until these are collocated.
	$S_1$ already contains a ground set of size $k-1$ (that it cannot split), hence \ALG must collocate $\{x_0, y_0\}$ in another cluster.

	We continue to construct ground sets containing the current alone node from $S$.
	To preserve a feasible partition of nodes into clusters after collocating $\{x_0, y_0\}$,
	\ALG must replace $x_0$ with another single node (that we name $x_1$).
	Then, we merge the ground sets of $x_1$ and the largest ground set $C$ s.t. and $C \neq \{x_0,y_0\}$~$I(C) = I(x_1)$.
	Continuing this, we observe a sequence $X := x_0, x_1, x_2\dots$ of alone nodes that visit $S$.
	(Note that subsequent ground sets are different than the first ground set $\{x_0, y_0\}$, whose nodes were initially in different clusters.)

	We claim that this procedure (of constructing ground sets) can continue as long as a perfect partition of ground sets still exists.
	%and we construct a sequence of ground sets $X := \langle \{x_0, y_0\}, \{x_1, y_2\}, \{x_2, y_2\}, \ldots \rangle$ of length $\Omega(k \cdot \ell)$.
	Consider the moment of creating a ground set $\{x_i, y_i\}$.
	If a perfect partition of ground sets still exists, then
	\ALG{} must collocate $\{x_i,y_i\}$ in a~cluster other than $S$.
	To this end, it replaces $x_i$ with another alone node $x_{i+1}$.
	To continue, we must guarantee that
	\maciek{elaborate}

	We end this procedure prematurely (before all nodes visit $S$) to keep \OPT's cost low.
	The sequence must end when no perfect partition of ground sets exists ---
	however, we end it when \emph{two more} ground sets are still possible to give.
	Precisely, we repeat constructing ground sets,
	until there are only two alone nodes left that originate from the same cluster.
	We denote these two nodes by $x^*, y^*$.
	Observe that at this point there are at most $\ell+1$ alone nodes left:
	otherwise, by pigeon hole principle, there would be more pairs of alone nodes that were initially in the same cluster.
